---
title: "Transcriptional profile of Aspergillus fumigatus under azole treatment"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: simplex
    css: doc.css
---
```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE)
library(knitr)
library(tidyverse)
```

## 1 METHODS {.tabset}

### Workflow

### Libraries

```{r, eval = FALSE}
library(tidyverse)
library(ComplexHeatmap)
library(circlize)
library(GSEABase)
library(waRRior)
library(clusterProfiler)
```

### Functions

```{r, eval = FALSE}
# Genes with expression > 0 across samples
universe <- counts %>%
  as.data.frame() %>% 
  rownames_to_column("gene_id") %>% 
  mutate(total = rowSums(across(where(is.numeric)))) %>% 
  filter(total > 0) %>% 
  pull(gene_id)

go.path <- "ref_files/FungiDB-55_AfumigatusAf293_GO.gaf"

# GO annotation from Fungi DB
go.fungi.DB <- read.delim2(file   =  go.path, 
                           skip   = 1, 
                           sep    = "\t", 
                           header = FALSE) %>%
  as_tibble() %>% 
  dplyr::rename(gene_id = 2, go = 5, onthology = 9, term = 10) %>% 
  dplyr::select(go, gene_id, onthology, term)

go_enrichment <- function(genes2go, go, universe){
  
  genes.enrich <- enricher(gene          = genes2go,
                           pvalueCutoff  = 0.05,
                           universe      = universe,
                           pAdjustMethod = "BH",
                           TERM2GENE     = go)
  return(genes.enrich@result)
}

run_DESeq <- function(full.form, 
                      reduc.form, 
                      counts, 
                      meta.data, 
                      evo,
                      rev) {
  
  # Not normalized data
  dds <- DESeq2::DESeqDataSetFromMatrix(countData = counts,
                                        colData   = meta.data,
                                        design    = full.form)
  if (evo == TRUE & rev == FALSE) {
    #Set ct as control
    dds$str_res = relevel(dds$str_res, "ct_evo")
    #Set 0h as control
    dds$timePoint = relevel(dds$timePoint, "0h")
  } else if (evo == FALSE & rev == TRUE) {
    #Set ct as control
    dds$str_res = relevel(dds$str_res, "ct_rev")
    #Set 0h as control
    dds$timePoint = relevel(dds$timePoint, "0h")
  } else {
    dds$strain = relevel(dds$strain, "ct")
    #Set 0h as control
    dds$timePoint = relevel(dds$timePoint, "0h")
    #Set rev as control
    dds$resist = relevel(dds$resist, "rev")
  }

  dds.lrt <- DESeq(dds,
                   test    = "LRT",
                   reduced = reduc.form)
  
  res <- results(dds.lrt)
  
  return(res)
}
```

### Data
```{r, eval = FALSE}
# Load count table from Geo2RNAseq
load("/sbidata/apereira/analysis/rnaseq_phillip_EvoRev/src/rdata/R_workspace.RData")

# Importing file names
files.path <- "/sbidata/abarber/azole_evo/02_rnaseq/01_raw_fastq"
tbl4samples <- tibble(name = list.files(files.path,
                                        full.names = FALSE,
                                        pattern    = "fq")) 
# Metadata
coldata <- tbl4samples %>%
  separate(name, c("strain", "resist", "timePoint", "replicate", "pairEnd"),
           sep    = "_",
           remove = FALSE) %>%
  filter(pairEnd == "1.fq.gz") %>% 
  mutate(samples = paste(strain, resist, timePoint, replicate, sep = "_")) %>%
  mutate(str_res = paste(strain, resist, sep = "_")) %>% 
  mutate(str_res_tim = paste(strain, resist, timePoint, sep = "_")) %>% 
  rowid_to_column("index") %>% 
  mutate_if(is_character, as_factor)

#Filter evolved samples
evo.index <- coldata %>%
  filter(pairEnd == "1.fq.gz") %>% 
  filter(resist == "evo") %>% 
  pull(index)

#Filter reverted samples
rev.index <- coldata %>%
  filter(pairEnd == "1.fq.gz") %>% 
  filter(resist == "rev") %>% 
  pull(index)

# Import annotation
ref_files  <- "/sbidata/apereira/analysis/rnaseq_phillip_EvoRev/ref_files/"
annot      <- paste(ref_files, "afu_af293.annot", sep = "")

# Annotation for rRNA removal
afu_annot <- read_tsv(annot) %>% 
  dplyr::rename(description = 3, gene_id = 1) %>% 
  select(gene_id, description)

# Annotate counts
annotate.counts <- count_res$counts %>%
  as_tibble() %>%
  mutate(gene_id = rownames(count_res$counts)) %>%
  left_join(afu_annot, by = "gene_id") %>%
  select(gene_id, description, everything())

# Filtering rRNA genes
annotate.counts.no.rRNA <-  annotate.counts %>% 
  filter(!description == "rRNA") %>% 
  column_to_rownames("gene_id") %>% 
  select(-c("description"))

# Normalization
counts.no.rRNA.mrn <- get_mrn(annotate.counts.no.rRNA)

# Getting replicate mean
counts.no.rRNA.mrn.mean <-  counts.no.rRNA.mrn %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("samples") %>%
  left_join(coldata %>% select(samples, strain, resist, timePoint)) %>%
  column_to_rownames("samples") %>%
  group_by(strain, resist, timePoint) %>%
  nest() %>%
  mutate(data = map(data, ~ .x %>%
                      summarise(across(where(is.numeric), mean)))) %>%
  unnest(data)
```

### DEG LRT

```{r, eval = FALSE}

#Full formula
full <- ~ (strain + resist + timePoint)^2

#Variables to remove
to.remove <- ~ (strain + resist + timePoint + strain:resist + resist:timePoint)

#Run LRT test
res.evo.rev.straintime <- run_DESeq(full.form      =  full, 
                                    reduc.form     =  to.remove, 
                                    counts         =  counts, 
                                    meta.data      =  coldata,
                                    evo            = TRUE,
                                    rev            = TRUE)
```

### Heatmap

```{r, eval = FALSE}
# Expression data from 4h and 1h
# Sig genes from LRT test

counts.4h.1h <- counts.no.rRNA.mrn.mean.t %>%
  as.data.frame() %>%
  filter(gene_id %in% evo.rev.genes) %>%
  dplyr::mutate(total = rowSums(across(where(is.numeric)))) %>%
  filter(total > 0, !is.na(across(where(is.numeric)))) %>%
  dplyr::select(-total) %>% 
  column_to_rownames("gene_id") %>%
  dplyr::select(contains(c("4h", "1h")))

# Zscaling
counts.4h.1h.scaled <- t(scale(t(counts.4h.1h), center = TRUE)) %>% 
  as.data.frame() %>% 
  drop_na() %>% 
  as.matrix()

col_fun = colorRamp2(c(-2, -1, -0.5, 0, 0.5, 1, 2), 
                     col_fun(c(-2, -1, -0.5, 0, 0.5, 1, 2)))

heat.4h.1h <- Heatmap(matrix                   =  counts.4h.1h.scaled,
                   clustering_method_rows   = "complete",
                   clustering_method_columns = "complete",
                   clustering_distance_rows = "euclidean",
                   color_space              = col_fun,
                   cluster_rows             = TRUE,
                   cluster_columns          = TRUE,
                   row_dend_reorder         = TRUE,
                   column_dend_reorder      = TRUE,
                   row_km                   = 4,
                   # column_km                = 4,
                   row_gap                  = unit(0.02, "npc", data=NULL),
                   show_row_names           = F,
                   name                     = "Z-score row scaled")

#Plot Heatmap
pdf(paste(path, "heat.1h.4h.pdf", sep = ""),
    width = 12,
    height = 12)
print(heat.4h.1h)
dev.off()

#Annotate Clusters
heat.4h.1h.clusters <- heatmap_extract_cluster(heatmap_obj = heat.4h.1h, 
                                            matrix_obj  = counts.4h.1h.scaled, 
                                            which       = "row")

# Gene enrichment analysis

root    <- "/sbidata/apereira/analysis/rnaseq_phillip_EvoRev/"
goSlim.path <- "ref_files/goslim_aspergillus.obo"
goslim <- getOBOCollection(goSlim.path)


heat.4h.1h.clusters.annot <- heat.4h.1h.clusters %>% 
  as.data.frame() %>% 
  dplyr::rename(gene_id = ID, cluster = Cluster) %>%
  group_by(cluster) %>% 
  nest() %>%
  mutate(enrich.clus = map(data, ~ go_enrichment(genes2go  = .x %>% pull(gene_id), 
                                                 go        = go.fungi.DB,
                                                 universe  = universe) %>%
                              filter(p.adjust < 0.05, qvalue < 0.05))) %>%
  dplyr::mutate(goColl = map(enrich.clus, ~ GOCollection(.x %>% 
                                                    mutate_if(is.factor, as.character) %>%
                                                    drop_na() %>% 
                                                    pull(ID)))) %>% 
  dplyr::mutate(BP = map(goColl, ~ GSEABase::goSlim(.x, goslim, "BP") %>% 
                              filter(Count > 0) %>% 
                              dplyr::select(Term))) %>% 
  dplyr::mutate(MF = map(goColl, ~ GSEABase::goSlim(.x, goslim, "MF") %>% 
                              filter(Count > 0) %>% 
                              dplyr::select(Term))) %>% 
  dplyr::mutate(CC = map(goColl, ~ GSEABase::goSlim(.x, goslim, "CC") %>% 
                              filter(Count > 0) %>% 
                              dplyr::select(Term))) %>%
  dplyr::select(-data, -goColl, -enrich.clus)  %>%
  pivot_longer(-cluster, names_to = "Onthology", values_to = "value") %>% 
  unnest(value) %>% 
  arrange(cluster)

saveRDS(heat.4h.1h.clusters.annot, "src/deg_time_course_2/tbls/heatmap1h.4h.go.annot.rds")
  

  
# 2. Heatmap 1765 sig. genes for 1h 
################################################################################
# Filtering sig. genes 
sig.genes.1h <- all.evo.logf %>% 
  filter(time == "1h" & padj < 0.05) %>% 
  pull(genes) %>% 
  unique()

# Filtering samples
counts.sig.1h <- counts.no.rRNA.mrn.mean.t %>%
  as.data.frame() %>%
  filter(gene_id %in% sig.genes.1h) %>%
  dplyr::mutate(total = rowSums(across(where(is.numeric)))) %>%
  filter(total > 0, !is.na(across(where(is.numeric)))) %>%
  dplyr::select(-total) %>% 
  column_to_rownames("gene_id") %>%
  dplyr::select(contains("1h"))

# Zscaling
counts.sig.1h.scaled <- t(scale(t(counts.sig.1h), center = TRUE)) %>% 
  as.data.frame() %>% 
  drop_na() %>% 
  as.matrix()

col_fun = colorRamp2(c(-2, -1, -0.5, 0, 0.5, 1, 2), 
                     col_fun(c(-2, -1, -0.5, 0, 0.5, 1, 2)))

heat.sig.1h <- Heatmap(matrix                   =  counts.sig.1h.scaled,
                      clustering_method_rows    = "complete",
                      clustering_method_columns = "complete",
                      clustering_distance_rows = "euclidean",
                      color_space              = col_fun,
                      cluster_rows             = TRUE,
                      cluster_columns          = TRUE,
                      row_dend_reorder         = TRUE,
                      column_dend_reorder      = TRUE,
                      row_km                   = 3,
                      # column_km                = 4,
                      row_gap                  = unit(0.02, "npc", data=NULL),
                      show_row_names           = F,
                      name                     = "Z-score row scaled")


#Plot Heatmap
pdf(paste(path, "heat.sig.1h.pdf", sep = ""),
    width = 12,
    height = 12)
print(heat.sig.1h)
dev.off()

#Annotate Clusters
heat.sig.1h.clusters <- heatmap_extract_cluster(heatmap_obj = heat.sig.1h, 
                                               matrix_obj  = counts.sig.1h.scaled, 
                                               which       = "row")

# Gene enrichment analysis

heat.sig.1h.clusters.annot <- heat.sig.1h.clusters %>% 
  as.data.frame() %>% 
  dplyr::rename(gene_id = ID, cluster = Cluster) %>%
  group_by(cluster) %>% 
  nest() %>%
  mutate(enrich.clus = map(data, ~ go_enrichment(genes2go  = .x %>% pull(gene_id), 
                                                 go        = go.fungi.DB,
                                                 universe  = universe) %>%
                             filter(p.adjust < 0.05, qvalue < 0.05))) %>%
  dplyr::mutate(goColl = map(enrich.clus, ~ GOCollection(.x %>% 
                                                           mutate_if(is.factor, as.character) %>%
                                                           drop_na() %>% 
                                                           pull(ID)))) %>% 
  dplyr::mutate(BP = map(goColl, ~ GSEABase::goSlim(.x, goslim, "BP") %>% 
                           filter(Count > 0) %>% 
                           dplyr::select(Term))) %>% 
  dplyr::mutate(MF = map(goColl, ~ GSEABase::goSlim(.x, goslim, "MF") %>% 
                           filter(Count > 0) %>% 
                           dplyr::select(Term))) %>% 
  dplyr::mutate(CC = map(goColl, ~ GSEABase::goSlim(.x, goslim, "CC") %>% 
                           filter(Count > 0) %>% 
                           dplyr::select(Term))) %>%
  dplyr::select(-data, -goColl, -enrich.clus)  %>%
  pivot_longer(-cluster, names_to = "Onthology", values_to = "value") %>% 
  unnest(value) %>% 
  arrange(cluster)

# 3. Heatmap 4224 sig. genes for 4h 
################################################################################
# Filtering sig. genes 
sig.genes.4h <- all.evo.logf %>% 
  filter(time == "4h" & padj < 0.05) %>% 
  pull(genes) %>% 
  unique()

# Filtering samples
counts.sig.4h <- counts.no.rRNA.mrn.mean.t %>%
  as.data.frame() %>%
  filter(gene_id %in% sig.genes.4h) %>%
  dplyr::mutate(total = rowSums(across(where(is.numeric)))) %>%
  filter(total > 0, !is.na(across(where(is.numeric)))) %>%
  dplyr::select(-total) %>% 
  column_to_rownames("gene_id") %>%
  dplyr::select(contains("4h"))

# Zscaling
counts.sig.4h.scaled <- t(scale(t(counts.sig.4h), center = TRUE)) %>% 
  as.data.frame() %>% 
  drop_na() %>% 
  as.matrix()

col_fun = colorRamp2(c(-2, -1, -0.5, 0, 0.5, 1, 2), 
                     col_fun(c(-2, -1, -0.5, 0, 0.5, 1, 2)))

heat.sig.4h <- Heatmap(matrix                   =  counts.sig.4h.scaled,
                       clustering_method_rows    = "complete",
                       clustering_method_columns = "complete",
                       clustering_distance_rows = "euclidean",
                       color_space              = col_fun,
                       cluster_rows             = TRUE,
                       cluster_columns          = TRUE,
                       row_dend_reorder         = TRUE,
                       column_dend_reorder      = TRUE,
                       row_km                   = 5,
                       # column_km                = 4,
                       row_gap                  = unit(0.02, "npc", data=NULL),
                       show_row_names           = F,
                       name                     = "Z-score row scaled")


#Plot Heatmap
pdf(paste(path, "heat.sig.4h.pdf", sep = ""),
    width = 12,
    height = 12)
print(heat.sig.4h)
dev.off()

#Annotate Clusters
heat.sig.4h.clusters <- heatmap_extract_cluster(heatmap_obj = heat.sig.4h, 
                                                matrix_obj  = counts.sig.4h.scaled, 
                                                which       = "row")

# Gene enrichment analysis

heat.sig.4h.clusters.annot <- heat.sig.4h.clusters %>% 
  as.data.frame() %>% 
  dplyr::rename(gene_id = ID, cluster = Cluster) %>%
  group_by(cluster) %>% 
  nest() %>%
  mutate(enrich.clus = map(data, ~ go_enrichment(genes2go  = .x %>% pull(gene_id), 
                                                 go        = go.fungi.DB,
                                                 universe  = universe) %>%
                             filter(p.adjust < 0.05, qvalue < 0.05))) %>%
  dplyr::mutate(goColl = map(enrich.clus, ~ GOCollection(.x %>% 
                                                           mutate_if(is.factor, as.character) %>%
                                                           drop_na() %>% 
                                                           pull(ID)))) %>% 
  dplyr::mutate(BP = map(goColl, ~ GSEABase::goSlim(.x, goslim, "BP") %>% 
                           filter(Count > 0) %>% 
                           dplyr::select(Term))) %>% 
  dplyr::mutate(MF = map(goColl, ~ GSEABase::goSlim(.x, goslim, "MF") %>% 
                           filter(Count > 0) %>% 
                           dplyr::select(Term))) %>% 
  dplyr::mutate(CC = map(goColl, ~ GSEABase::goSlim(.x, goslim, "CC") %>% 
                           filter(Count > 0) %>% 
                           dplyr::select(Term))) %>%
  dplyr::select(-data, -goColl, -enrich.clus)  %>%
  pivot_longer(-cluster, names_to = "Onthology", values_to = "value") %>% 
  unnest(value) %>% 
  arrange(cluster)
```

### DEG Wald and ORA
```{r, eval = FALSE}
dds <- DESeq2::DESeqDataSetFromMatrix(countData = counts,
                                      colData   = coldata,
                                      design    = ~ str_res_tim)

dds$str_res_tim = relevel(dds$str_res_tim, "ct_evo_0h")

# Parallelization
bpparam <- MulticoreParam(workers = 16)

dds <- DESeq(dds, 
             parallel = TRUE, 
             BPPARAM  = bpparam)

# Contrasting evo x ct and rev x ct
# Not p-value filtering
contrasts.all <- coldata %>%
  filter(replicate == 1, !strain == "ct", resist == "evo") %>% 
  select(str_res_tim) %>% 
  mutate(str_res_tim2 = str_res_tim) %>% 
  mutate(str_res_tim2 = str_replace(str_res_tim2, "^[abd]", "ct")) %>% 
  mutate(comp3 = as.character(str_res_tim)) %>% 
  mutate(comp4 = str_res_tim2) %>%
  mutate(comp3 = str_replace(comp3, "evo", "rev")) %>% 
  mutate(comp4 = str_replace(comp4, "evo", "rev")) %>% 
  dplyr::rename(comp1 = 1, comp2 = 2) %>% 
  mutate(comp1 = as.character(comp1)) %>%
  mutate(deseq.evo = map2(comp1, comp2, ~ results(dds, 
                                                  contrast = c("str_res_tim", 
                                                               as.character(.x), 
                                                               as.character(.y)),
                                                  parallel = TRUE,
                                                  BPPARAM  = bpparam) %>% 
                          as.data.frame())) %>% 
                          # filter(padj < 0.05) %>% 
  mutate(deseq.rev = map2(comp3, comp4, ~ results(dds, 
                                                  contrast = c("str_res_tim", 
                                                               as.character(.x), 
                                                               as.character(.y)),
                                                  parallel = TRUE,
                                                  BPPARAM  = bpparam) %>% 
                            as.data.frame()))

all.evo.logf <- contrasts.all %>%
  group_by(comp1) %>% 
  mutate(deseq.evo.logf = map(deseq.evo, ~ .x %>% 
                                as.data.frame() %>% 
                                rownames_to_column("genes")))  %>%
  select(deseq.evo.logf) %>%
  unnest(deseq.evo.logf) %>% 
  dplyr::rename(sample = comp1) %>% 
  separate(sample, c("str", "resis", "time"), remove = FALSE)


all.rev.logf <- contrasts.all %>%
  group_by(comp3) %>% 
  mutate(deseq.rev.logf = map(deseq.rev, ~ .x %>% 
                                as.data.frame() %>% 
                                rownames_to_column("genes")))  %>%
  select(deseq.rev.logf) %>%
  unnest(deseq.rev.logf) %>% 
  dplyr::rename(sample = comp3) %>% 
  separate(sample, c("str", "resis", "time"), remove = FALSE)

all.logf <- bind_rows(all.evo.logf, all.rev.logf) 

```

### Other plots
```{r, eval = FALSE}

# Methyltrasnferase activity
###############################################################################

all.logf %>%
  dplyr::rename(gene_id = genes) %>% 
  left_join(go.fungi.DB %>% distinct(gene_id, term)) %>% 
  dplyr::filter(padj < 0.05, str_detect(term, "histone methyltransferase")) %>%
  distinct() %>%
  group_by(str, resis, time) %>% 
  ggplot(aes(x = str, y = log2FoldChange)) +
  geom_boxplot(aes(fill = time), position = "dodge") +
  facet_wrap(~ resis) +
  scale_colour_brewer(name       = "Time Point", 
                      palette    = "Set1", 
                      aesthetics = c("fill")) +
  ylab("Log2FoldChange relative to control") +
  ggtitle("Histone methyltransferase activity") +
  xlab("Lineages") +
  theme_bw() +
  ggsave("src/deg_time_course_2/plots/hist.methyl.pdf", 
         device    = "pdf",
         width     = 8,
         height    = 8, 
         limitsize = FALSE)

all.logf %>%
  dplyr::rename(gene_id = genes) %>% 
  left_join(go.fungi.DB %>% distinct(gene_id, term)) %>% 
  dplyr::filter(padj < 0.05, str_detect(term, "transpos")) %>%
  distinct() %>%
  group_by(str, resis) %>% 
  ggplot(aes(x = term, y = log2FoldChange)) +
  geom_boxplot(aes(fill = time), 
               position = position_dodge(preserve = "single",
                                         width = 1)) +
  facet_wrap(facets =  ~ resis + str, 
             scales = "free_x") +
  scale_colour_brewer(name       = "Time Point", 
                      palette    = "Set1", 
                      aesthetics = c("fill")) +
  ylab("Log2FoldChange relative to control") +
  ggtitle("Transposon activity") +
  xlab("Transposons") +
  theme_bw() +
  coord_flip() +
  geom_hline(yintercept = 0) +
  ggsave("src/deg_time_course_2/plots/transp.pdf", 
         device    = "pdf",
         width     = 10,
         height    = 10, 
         limitsize = FALSE)



all.logf %>%
  dplyr::rename(gene_id = genes) %>% 
  left_join(go.fungi.DB %>% distinct(gene_id, term)) %>% 
  dplyr::filter(padj < 0.05, str_detect(term, "histone demethylase")) %>%
  distinct() %>%
  group_by(str, resis, time) %>% 
  ggplot(aes(x = str, y = log2FoldChange)) +
  geom_col(aes(fill = time), position = position_dodge(preserve = "total",
                                         width = 1)) +
  facet_wrap(~ resis) +
  scale_colour_brewer(name       = "Time Point", 
                      palette    = "Set1", 
                      aesthetics = c("fill")) +
  ylab("Log2FoldChange relative to control") +
  ggtitle("Histone demethylase activity") +
  xlab("Lineages") +
  theme_bw() +
  ggsave("src/deg_time_course_2/plots/hist.demethyl.pdf", 
         device    = "pdf",
         width     = 8,
         height    = 8, 
         limitsize = FALSE)



all.logf %>%
  dplyr::rename(gene_id = genes) %>% 
  left_join(go.fungi.DB %>% distinct(gene_id, term)) %>% 
  dplyr::filter(padj < 0.05, str_detect(term, c("histone acetyltransferase", 
                                                "histone deacetyl"))) %>%
  distinct() %>%
  group_by(str, resis, time) %>% 
  ggplot(aes(x = str, y = log2FoldChange)) +
  geom_boxplot(aes(fill = time), position = "dodge") +
  facet_wrap(~ resis)

all.logf %>%
  dplyr::rename(gene_id = genes) %>% 
  left_join(go.fungi.DB %>% distinct(gene_id, term)) %>% 
  dplyr::filter(padj < 0.05, str_detect(term, c("histone acetyltransferase", 
                                                "histone deacetyl"))) %>%
  distinct() %>%
  group_by(str, resis, time) %>% 
  ggplot(aes(x = str, y = log2FoldChange)) +
  geom_boxplot(aes(fill = time), position = "dodge") +
  facet_wrap(~ resis) +
  scale_colour_brewer(name       = "Time Point", 
                      palette    = "Set1", 
                      aesthetics = c("fill")) +
  ylab("Log2FoldChange relative to control") +
  ggtitle("Histone acetyltransferase activity") +
  xlab("Lineages") +
  theme_bw() +
  ggsave("src/deg_time_course_2/plots/hist.acethyl.pdf", 
         device    = "pdf",
         width     = 8,
         height    = 8, 
         limitsize = FALSE)

all.logf %>%
  dplyr::rename(gene_id = genes) %>% 
  left_join(go.fungi.DB %>% distinct(gene_id, term)) %>% 
  dplyr::filter(padj < 0.05, str_detect(term, c("transcription factor"))) %>%
  distinct() %>%
  group_by(str, resis, time) %>% 
  ggplot(aes(x = str, y = log2FoldChange)) +
  geom_boxplot(aes(fill = time), position = "dodge") +
  facet_wrap(~ resis) +
  scale_colour_brewer(name       = "Time Point", 
                      palette    = "Set1", 
                      aesthetics = c("fill")) +
  ylab("Log2FoldChange relative to control") +
  ggtitle("Transcription factor activity") +
  xlab("Lineages") +
  theme_bw() +
  ggsave("src/deg_time_course_2/plots/transc.pdf", 
         device    = "pdf",
         width     = 12,
         height    = 8, 
         limitsize = FALSE)
```

### ERGOSTEROL PATHWAY

```{r, eval = FALSE}
#Identify Genes based on Metabolic Pathway - FungiDB
# Ergosterol Biosynthesis II - Metacyc

erg.path <- read_tsv("ref_files/erg.path.tsv") %>% 
  as_tibble() %>% 
  dplyr::select(1,3,4) %>% 
  dplyr::rename(gene_id = 1,
                term    = 2,
                alias   = 3) %>% 
  mutate(alias = ifelse(alias == "N/A", map(gene_id, ~ .x), alias)) %>% 
  unnest_auto(alias)

str(erg.path)


tmp <- all.logf %>% 
  dplyr::filter(genes  %in% erg.path$gene_id, time == "4h") %>%
  left_join(erg.path %>% dplyr::rename(genes = gene_id)) %>%
  mutate(padj.cut = cut(padj, breaks   = c(0,0.01,0.05, 1), 
                        include.lowest = T, 
                        ordered_result = T), .before = 1) %>% 
  ggplot(aes(x = resis, 
             y = log2FoldChange)) +
  geom_line(aes(group = str, 
                color = str)) +
  scale_colour_brewer(name       = "Lineages", 
                      palette    = "Set1", 
                      aesthetics = c("color")) +
  new_scale_color() +
  geom_point(aes(color = padj.cut, 
                 size  = padj)) +
  scale_colour_manual(name   = "Adjusted p-value",
                      labels = c("[0-0.01]", "(0.01-0.05]", "(0.05-1]"),
                      values = c("#e66101","#fdb863", "#5e3c99"),
                      aesthetics = "color") +
  scale_size(range  = c(0.5,3),
             breaks = c(0,0.01,0.05, 1),
             labels = c("0","0.01","0.05", "1"),
             guide  = "legend",
             trans  = "reverse",
             name   = "Importance") +
  facet_wrap(facets =  ~ alias,
             scales = "free_y") +
  ylab("Log2FoldChange relative to control") +
  ggtitle("Ergosterol Superpathway II at time point 4h") +
  geom_hline( yintercept = 0) +
  xlab("Resistance") +
  theme_bw() +
  ggsave("src/deg_time_course_2/plots/erg.path.pdf", 
         device    = "pdf",
         width     = 20,
         height    = 20, 
         limitsize = FALSE)
```

## 2 RESULTS {.tabset}

### BACKGROUND

- Azoles exert their action by inhibiting the C14α demethylation of lanosterol in fungi
- Interference in the synthesis of ergosterol in the fungal cell membrane
- Azoles are fungistatic drugs

- Resistance levels:

| Lineage| Evolved    | Reverted   |
|--------|------------|------------|
|A       | 16 mg/L    | 12 mg/L    |
|B       | 10.67 mg/L | 5 mg/L     |
|D       | 12.67 mg/L | 5 mg/L     |
Ancestor ~1.8 mg/L |    |          |

### Heatmap 1

- Expression data from 1hr and 4hr
- 714 significative genes from LRT test  

- 4h and 1h clustered separately
- At 1h
    - Cluster 2 being expressed
    - Response to stress (BP)
    - DNA metabolic process (BP)
    - Vesicle-mediated transport (BP)
    - Protein catabolic process (BP)
    - Transport (BP)
    - Apparently no difference between EVO and REV
    
    - Hypothesis
        - Cells are growing and starting to generate an anti-stress response
        - DNA biosynthesis for supporting growth
        - Transport:
Filamentous fungi grow through polarized hyphal tip extension and septation, resulting in
elongated cellular compartments which are key to hyphal growth, development and survival
[1]. Filamentous growth pattern requires both short- and long-distance intracellular
transport, as well as spatiotemporal coordination of a multitude of processes to generate and
deliver nutrients and proteins [2]. Intracellular trafficking of essential components through
membrane-bound vesicles to the plasma membrane for exocytosis is important for cellular
growth and development. As such, cytoskeletal components, including the actin and
microtubule networks and the molecular motors that use these networks, are integral to
morphogenesis in filamentous fungi [2].
      - Protein Catabolism??

- At 4h
    - Cluster 4 upregualted only in the controls
        - lipid metabolic process (BP)
        - Peroxissome activity (CC)
        - Hypothesis:
            - Ergosterol pathway activated, but inhibited by drug
            - Controls are susceptible
            - Peroxissomes are important site for fatty acids β-oxidation.
    - Cluster 1 upregualted in treated group
        - Response to stress (BP)
        - Pathogenesis (BP)
        - Cellular homeostasis (BP)
        - Secondary metabolic process (BP)
            - Transporter activity (MF)
                - May include pumps
            - Transferase activity (MF): Translocation of functional groups
                - May include methyl -> Epigenetic mark
                - Methylation typically acts to repress gene transcription.
            - Hydrolase activity (MF): Covalent bond breaker
        - Hypothesis:
            - Secondary metabolite production in response to stress, what leads to pathogenesis
            - Secondary metabolite produced by transferases and hydrolases
            - SM secreation through pumps.
            

```{r out.height = "1000px", out.width='1200px', echo=F}
knitr::include_graphics("../src/deg_time_course_2/plots/heat.1h.4h.pdf")
```

- GO Slim annotation

```{r, echo = FALSE}
tbl1 <- read_rds("../src/deg_time_course_2/tbls/heatmap1h.4h.go.annot.rds")
kable(x         = tbl1, 
      digits    = 2, 
      longtable = TRUE,
      align     = "l")
```

### Heatmap 2

- Only genes significant in any strain at 1 hr
- 1765 significative genes
- P-value < 0.05
- Expression data from 1hr

- Cluster 2 upregulated in treated ones
    - The same as cluster 1 on previous heatmap
- Cluster 3 upregulated in D Evo
    - Hypothesis:
        - Cell wall remodelling

```{r out.height = "1000px", out.width='1200px', echo=F}
knitr::include_graphics("../src/deg_time_course_2/plots/heat.sig.1h.pdf")
```

- GO Slim annotation

```{r, echo = FALSE}
tbl1 <- read_rds("../src/deg_time_course_2/tbls/heat.sig.1h.go.annot.rds")
kable(x         = tbl1, 
      digits    = 2, 
      longtable = TRUE,
      align     = "l")
```

### Heatmap 3

- Only genes significant in any strain at 4 hr
- 4224 significative genes
- P-value < 0.05
- Expression data from 4hr

- Cluster 2 Upregulated in B Evo and less in B rev
    - Hypothesis:
        - Transcription factors being activated and regulating metabolism  and organelles
        
- Cluster 1 upregulated in D EVO
    - Envolved transcription factor activity
    
- Cluster 4 upregulated in D REV
    - Different set of genes being affected by transcription factors
    
- Cluster 5 not important for resistance

- Cluster 3 cannot define uniquely any lineage.

```{r out.height = "1000px", out.width='1200px', echo=F}
knitr::include_graphics("../src/deg_time_course_2/plots/heat.sig.4h.pdf")
```

- GO Slim annotation

```{r, echo = FALSE}
tbl1 <- read_rds("../src/deg_time_course_2/tbls/heat.sig.4h.go.annot.rds")
kable(x         = tbl1, 
      digits    = 2, 
      longtable = TRUE,
      align     = "l")
```

### Acetyl

```{r out.height = "1000px", out.width='1200px', echo=F}
knitr::include_graphics("../src/deg_time_course_2/plots/hist.acethyl.pdf")
```

### methylT

```{r out.height = "1000px", out.width='1200px', echo=F}
knitr::include_graphics("../src/deg_time_course_2/plots/hist.methyl.pdf")
```

### demethyl

```{r out.height = "1000px", out.width='1200px', echo=F}
knitr::include_graphics("../src/deg_time_course_2/plots/hist.demethyl.pdf")
```

### Transp

```{r out.height = "1000px", out.width='1200px', echo=F}
knitr::include_graphics("../src/deg_time_course_2/plots/transp.pdf")
```

### Transc

```{r out.height = "1000px", out.width='1200px', echo=F}
knitr::include_graphics("../src/deg_time_course_2/plots/transc.pdf")
```

### ERG

```{r out.height = "1000px", out.width='1200px', echo=F}
knitr::include_graphics("../src/deg_time_course_2/plots/erg.path.pdf")
```